`make.modifiers`

### 2023-02-22 inst/gitbook/03_make.modifiers.Rmd
added the modifiers bits

# Making modifiers with `make.modifiers()` {#makemodifiers}

@@

```{r}
    always.false <- function() return(FALSE)
    always.one <- function(x = NULL) return(1)

    test <- make.modifiers(branch.length = branch.length.trait,
                           speciation    = speciation.trait,
                           condition     = always.false, 
                           modify        = always.one)
```

## Combining and editing modifiers (`add`)

## Testing modifiers (`test`)

>>>>

# Modifying the birth-death process {#makemodifiers}

@@

For example in the `speciation` part of a modifier, the default is to trigger an event from a uniform distribution and then check if that value is smaller than ($$speciation/(speciation + extinction)$$):

```{r}
## The default speciation algorithm
speciation <- function(bd.params) {
    ## Randomly trigger an event
    trigger_event <- runif(1)

    ## Speciate?
    return(trigger_event < (bd.params$speciation/(bd.params$speciation + bd.params$extinction)))
}
```

It is possible with some specific condition to modify this trigger by providing a `condition` and `modify` function to the speciation function.
That is, if the `condition` is met, apply the `modify` function to the algorithm.
For example here we can edit the default speciation function to have a modification (`modify = double.the.trigger`: doubling the trigger value) happening half the time (`condition = half.the.time`):

```{r}
## A conditional function that triggers half the time
half.the.time <- function() return(sample(c(TRUE, FALSE), 1))
## A modification that doubles the value to trigger the event
double.the.trigger <- function(x) return(x*2)
## A conditional modifier
make.modifiers(speciation = speciation,
               condition  = half.the.time, 
               modify     = double.the.trigger)
```
Effectively, this will internally modify the speciation function as follows

```{r, eval = FALSE}
## The default speciation algorithm
speciation <- function(bd.params) {
    ## Randomly trigger an event
    trigger_event <- runif(1)

    ## Modify the triggering
    if(half.the.time()) { ## This running the half.the.time function drawing TRUE or FALSE randomly
        trigger_event <- double.the.trigger(trigger_event) ## This will double the random value trigger_event
    }

    ## Speciate?
    return(trigger_event < (bd.params$speciation/(bd.params$speciation + bd.params$extinction)))
}
```

These `condition` and `modify` functions can be applied to all the `modifiers` elements (`selection`, `branch.length` and `speciation`).
However, they are typically used in `events` that will modify... the `modifiers` ([see the `events` section](#makeevents))!
